---
title: "Walkthrough for Chapter 5- Data Wrangling"
subtitle: ""
output:
  github_document:
    toc: true
allow_html: true
editor_options:
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(nycflights13)

theme_set(theme_light())

knitr::opts_chunk$set(
  cache = FALSE,
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```

# Intro

Using `nycflights13::flights` dataset.
```{r}

flights %>% 
  head(5)

```

# filter()

Use an `&` filter by just adding multiple conditions.
```{r}

flights %>% 
  filter(month == 1, day == 1)

```

Have to assign to actually save the new data frame
```{r}

dec_25 <- flights %>% filter(month == 12, day == 25)

```

The comparisons to use are `>`, `>=`, `<`, `<=`, `!=`, and `==`

Logical operators like `&` and `|`, and `%in%` are also used.
```{r}

flights %>% 
  filter(month %in% c(11, 12))

```

When an `NA` value is involved in a filter, the result will almost always be `NA` as well
```{r}

NA > 5

```

```{r}

NA + 10

```

These operations make sense when we consider adding a known and an unknown; the result will also be unknown!

## Exercises

1. Find all flights that

- Had an arrival delay of two or more hours
```{r}

flights %>% 
  filter(arr_delay <= 120)

```

- Flew to Houston (IAH or HOU)
```{r}

flights %>% 
  filter(dest %in% c("IAH", "HOU"))

```

- Departed between midnight and 6am (inclusive)
```{r}

flights %>% 
  filter(dep_time >= 0 & dep_time <= 600)

```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}

flights %>% 
  filter(between(dep_time, 0, 600))

```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}

flights %>% 
  filter(is.na(dep_time))

```

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}

NA ^ 0 # anything to the 0th power equals 1

```

```{r}

NA | TRUE # anything or TRUE has to be TRUE

```

```{r}

FALSE & NA # anything and false has to be FALSE

```

# arrange()

`arrange()` allows us to order rows based on values.

```{r}

flights %>% 
  arrange(year)

```

```{r}

flights %>% 
  arrange(year, month, day)

```

Use `desc()` to sort in...descending order!
```{r}

flights %>% 
  arrange(desc(month))

```

Missing values are always sorted to the end
```{r}

tibble(x = c(1:5, NA)) %>% 
  arrange(x)

```

## Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}

tibble(x = c(1:5, NA, 6, NA)) %>% 
  arrange(desc(is.na(x)))

```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}

flights %>% 
  arrange(desc(dep_delay)) %>% 
  select(dep_delay)

```

# select()

Select columns!
```{r}

flights %>% 
  select(year, month, day)

```

Use helper functions like:

- `starts_with()`
- `ends_with()`
- `contains()`
- `matches()`- reg expressions
- `num_range()`- `num_range("x", 1:3)` would reurn x1, x2, x3

`rename()` is a variant of select and returns all columns
```{r}

flights %>% 
  rename(Year = year)

```

`everything()` is a nice select helper, especially when you want to move a few columns to the front
```{r}

flights %>% 
  select(origin, everything())

```

## Exercises



# mutate()


# summarise()


# grouped mutate (and filters)
