---
title: "Walkthrough for chapter 10- Tibbles"
subtitle: ""
output:
  github_document:
    toc: true
allow_html: true
editor_options:
  chunk_output_type: console
fig.width: 10
fig.height: 7
---

```{r setup}

library(tidyverse)
library(here)
library(knitr)

theme_set(theme_light())

opts_chunk$set(
  cache = TRUE,
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```

----------

# Creating tibbles

Tibbles are essentially data frames with a few nicer functionalities

as_tibble() is the function that coerces a data frame to a tibble
```{r}

as_tibble(iris)

```

tibble() can be used to create a new data object type tibble
```{r}

tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)

```

tibble() does not change the input type (e.g. string --> factor), doesn't change the variable names, and doesn't create row names.

Columns can also be non-syntatic
```{r}

tibble(
  `:)` = 1:100,
  `:(` = 2
)

```

tribble() is another way to create a tibble and allows for traditional data entry to create a dataset
```{r}

tribble(
  ~x, ~y, ~z,
  
  "a", 2, 3.6,
  "b", 1, 8.5
)

```

----------

# Tibbles vs data.frame

There are two main differences between tibbles and data frames:

- **Printing**

Default prints to 10 rows
```{r}

as_tibble(iris) %>% print()

iris %>% print()

```

The width = inf option allows to print all columns
```{r}

print(as_tibble(mpg), n = 15, width = Inf)

```

View() is also great for easily viewing a dataset in the RStudio data viewer
```{r}

iris %>% View()

```

- **Subsetting**

Extract a column by name using [[]] or $
```{r}

df <- tibble(
  y = 1:10,
  x = letters[1:10],
  z = rnorm(n = 10)
)

df$x

df[["x"]]

```

And in a pipe, . is the data set placeholder
```{r}

df %>% 
  filter(z < 0) %>% 
  .$z 

```

----------

# Exercises



**1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).**
```{r}

print(mtcars)

print(as_tibble(mtcars))

```

Tibble version prints a clean 10 rows only and...it says it's a tibble lol.


**2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?**

```{r}    
df <- data.frame(abc = 1, xyz = "a")

str(df)

df$x
str(df[, "xyz"])
str(df[, c("abc", "xyz")])

tib <- tibble(
  abc = 1,
  xyz = "a"
)

str(tib)

tib$x
tib[, "xyz"]
tib[, c("abc", "xyz")]

```

Advantages of the tibble:

- Does not coerce column xyz to be factor instead of chr (and therefore can't access x through $, a column selector)
- The bracket subset returns type factor for a single column subset, type data frame for the multi-column. The tibble subset always returns a tibble.

**3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?**
```{r}

var <- "model"

mpg %>% 
  select(var)

```

**4. Practice referring to non-syntactic names in the following data frame by:**

```{r}

annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
  )

```

- Extracting the variable called 1.
```{r}

annoying %>% select(`1`)

```

- Plotting a scatterplot of 1 vs 2.
```{r}

annoying %>% 
  ggplot(aes(`1`, `2`)) +
  geom_point()

```

- Creating a new column called 3 which is 2 divided by 1.
```{r}

annoying <- annoying %>% 
  mutate(`3` = `2` / `1`)

```

- Renaming the columns to one, two and three.
```{r}

colnames(annoying)

annoying %>% 
  rename(
    one = `1`,
    two = `2`,
    three = `3`
  )

```

**5. What does tibble::enframe() do? When might you use it?**
```{r}

?enframe

```

enframe() & deframe() do converstions:

- **Vectors to data frames**
```{r}

1:5 %>% enframe()

```

- **Data frames to vectors**
```{r}

tibble(a = 1:5, b = letters[1:5]) %>% deframe()

```

6. What option controls how many additional column names are printed at the footer of a tibble?
```{r}

as_tibble(mpg) %>% print()

```

