---
title: "Self-learning for Chapter 3- Data Visualization"
output: 
  github_document:
    toc: true
editor_options:
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(knitr)

theme_set(theme_light())

opts_chunk$set(
  cache = TRUE,
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```


```{r, message=FALSE}

air_incidents <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv") 

```

# Basic scatterplot
```{r}

air_incidents %>% 
  ggplot(aes(fatal_accidents_85_99, fatalities_85_99)) +
  geom_point()

```
  
# Plot an aesthetic corresponding to another variable.
```{r}

air_incidents %>% 
  ggplot(aes(fatal_accidents_85_99, fatalities_85_99, size = avail_seat_km_per_week)) +
  geom_point()

```
  
# Plot a manual aesthetic
```{r}

air_incidents %>% 
  ggplot(aes(fatal_accidents_85_99, fatalities_85_99, size = avail_seat_km_per_week)) +
  geom_point(color = "purple")

```

# Facet by a third variable
```{r}

air_incidents %>% 
  filter(airline %in% c("Southwest Airlines",
                        "Alaska Airlines*",
                        "American*")) %>% 
  ggplot(aes(avail_seat_km_per_week, fatalities_85_99)) +
  geom_point() +
  facet_wrap(~ airline)

```

# Use two different geom()'s
```{r}

air_incidents %>% 
  ggplot(aes(incidents_85_99, incidents_00_14)) +
  geom_point() +
  geom_smooth(se = FALSE)

```

# Use a geom() that uses stat. transformations
```{r}

air_incidents %>% 
  arrange(desc(avail_seat_km_per_week)) %>% 
  slice(1:5) %>% 
  ggplot(aes(airline, avail_seat_km_per_week)) +
  geom_col()

```

This dataset already had precomputed values, but could've used geom_bar() if it was in a more raw data-like form.

# Use a positional adjustment in geom_boxplot()
```{r}

air_incidents %>% 
  filter(airline %in% c("American*", "Southwest Airlines", "Air Canada")) %>% 
  ggplot(aes(airline, avail_seat_km_per_week)) +
  geom_boxplot(position = "dodge")

```

This would work if data was raw and each row was a flight with that flight's available seats

# Use coord_flip() and coord_polar()
```{r}

air_incidents %>% 
  filter(airline %in% c("American*", "Southwest Airlines", "Air Canada")) %>% 
  ggplot(aes(airline, avail_seat_km_per_week)) +
  geom_boxplot(position = "dodge") +
  coord_flip()

```





  