---
title: "Walkthrough for Chapter 3- Data Visualization"
output: 
  github_document:
    toc: true
editor_options:
  chunk_output_type: console
---


```{r}

library(tidyverse)

```

# 3.2- First steps
  
Basic scatter plot
```{r}

mpg %>% 
  ggplot() +
  geom_point(aes(displ, hwy))

```

## Exercises

1. Run ggplot(data = mpg). What do you see?
```{r}

mpg %>% 
  ggplot()

```
Empty plot template
  
2. How many rows are in mpg? How many columns?
```{r}

nrow(mpg)
ncol(mpg)
dim(mpg)

```
  
3. What does the drv variable describe? Read the help for ?mpg to find out.
```{r}
  
?mpg

```
  
4. Make a scatterplot of hwy vs cyl.
```{r}

mpg %>% 
  ggplot(aes(hwy, cyl)) +
  geom_point()

```
  
5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}

mpg %>% 
  ggplot(aes(class, drv)) +
  geom_point()

```

Uhhhh no.
  
----------
  
# 3.3- Aesthetic mappings

To assign an aesthetic to the points, include the aesthetic type inside aes()
```{r}

mpg %>% 
  ggplot() +
  geom_point(aes(displ, hwy, color = class))

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy, size = class)) +
  geom_point()

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy, alpha = class)) +
  geom_point()

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy, shape = class)) +
  geom_point()

```
  
To assign a manual aesthetic that doesn't convey info about a variable, place the aesthetic outside aes()
```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(color = "blue")

```
  
## Exercises

1. What’s gone wrong with this code? Why are the points not blue?
```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

```

The manually assigned aesthetic is inside aes() and so it's looking for variable information about "blue"
  
3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}

mpg %>%
  ggplot(aes(displ, hwy, color = cty)) +
  geom_point()

```

Color and shape are on a scale, shape can't be done.
  
4. What happens if you map the same variable to multiple aesthetics?
```{r}

mpg %>% 
  ggplot(aes(hwy, hwy)) +
  geom_point()

```

It works! But is useless.
  
5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}



```

  
6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}

mpg %>% 
  ggplot(aes(displ, hwy, color = displ < 5)) +
  geom_point()

```

Creates a binary TRUE/FALSE variable based on the condition provided.
  
----------
  
# 3.5- Facets
  
Facet by another variable with facet_wrap()
```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~ class, nrow = 2)

```
  
Facet by 2 variables with facet_grid()
```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point() +
  facet_grid(drv ~ cyl)

```

----------

# 3.6- Geometric objects

Geoms are used to place objects of different types onto a plot.

```{r}

mpg %>% 
  ggplot() +
  geom_point(aes(displ, hwy))

```
 
```{r}

mpg %>% 
  ggplot() +
  geom_smooth(aes(displ, hwy))

```

Change the object by adding an object characteristic (e.g. color, size, shape) inside aes() 

```{r}

mpg %>% 
  ggplot() +
  geom_smooth(aes(displ, hwy, linetype = drv))

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy, color = drv)) +
  geom_point() +
  geom_smooth(aes(linetype = drv))
  

```
 
Note that using aesthetics in ggplot() will apply those to every geom_*() function in the chain.

This insinuates that using different arguments in different geoms allows lots of flexibility
```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

```

## Exercises

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
```{r}

geom_line()
geom_boxplot()
geom_histogram()
geom_area()

```


2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
```

3. What does show.legend = FALSE do? 

Dunno, just use
```{r}
theme(legend.position = "none")
```

4. What does the se argument to geom_smooth() do?

Removes error bars

5. Will these two graphs look different? Why/why not?

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
  
Nope, inside ggplot() means inside all geom()'s  

6.Recreate the R code necessary to generate the following graphs.
```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point() +
  geom_smooth(aes(group = drv), se = FALSE)

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(se = FALSE)

```

```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = drv)) + 
  geom_smooth(aes(linetype = drv), se = FALSE)

```

----------

# 3.7- Statistical transformations

Some geom()'s compute statistics about the dataset behind the scenes. geom_bar() computes count by default, for instance.
```{r}

diamonds %>% 
  ggplot(aes(cut)) +
  geom_bar()

```

geom_bar() is associated with the stat stat_count() and these can be used interchangeably
```{r}

diamonds %>% 
  ggplot(aes(cut)) +
  stat_count()

```

Every geom has a default stat and every stat has a default geom. These usually don't need to be changed.

However, there are a few cases where we might want to change the stat:

- Overriding the default stat
```{r}

diamonds %>% 
  group_by(cut) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(cut, n)) +
  geom_bar(stat = "identity")

```

stat = "identity" is used when the y-axis value (like count, n) is explicitly in the dataset being plotted.

- Overriding the default mapping from transformed variables to aesthetics
```{r}

diamonds %>% 
  ggplot(aes(cut, ..prop.., group = 1)) +
  geom_bar()

```

- Drawing greater attention to the statistical tranformation in the code
```{r}

diamonds %>% 
  ggplot() +
  stat_summary(aes(cut, depth),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median
               )

```

## Exercises

1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

2. What does geom_col() do? How is it different to geom_bar()?

geom_bar() aggregates the dataset and computes a value, geom_col() is used when the aggregated value is already in the dataset

3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

4. What variables does stat_smooth() compute? What parameters control its behaviour?

```{r}

?stat_smooth

```

Computes a predicted value, lower/upper CI points around the mean, and a standard error.

5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))

```

```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))

```

In those plots, geom_bar() is aggregating the data by cut and then ..prop.. is calculating the proportion for each group *of that same group*. By adding group = 1, the proportions are calculated relative to the entire dataset instead of that specific group.

```{r}

diamonds %>% 
  ggplot() +
  geom_bar(aes(cut, ..prop.., group = 1))

```

----------

# 3.8- Position adjustments

Can use fill to color bars
```{r}

diamonds %>% 
  ggplot(aes(cut, fill = cut)) +
  geom_bar()

```

And a 2nd specified variable as the fill creates a stacked bar chart.
```{r}

diamonds %>% 
  ggplot(aes(cut, fill = clarity)) +
  geom_bar()

```

The position argument can be used to change from stacked bar chart to other styles
```{r}

diamonds %>% 
  ggplot(aes(cut, fill = clarity)) +
  geom_bar(alpha = 1/5, position = "identity")

diamonds %>% 
  ggplot(aes(cut, fill = clarity)) +
  geom_bar(position = "fill")

diamonds %>% 
  ggplot(aes(cut, fill = clarity)) +
  geom_bar(position = "dodge")

```

And position can also be useful in scatterplots; geom_jitter() is a common shorthand for geom_point(position = "jitter")
```{r}

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_jitter()

```

## Exercises

1. What is the problem with this plot? How could you improve it?

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()

```

The points overlap; we should use geom_jitter()
```{r}

mpg %>% 
  ggplot(aes(cty, hwy)) +
  geom_jitter()

```

2. What parameters to geom_jitter() control the amount of jittering?

```{r}

?geom_jitter

```

width and height control horizontal and vertical jitter.

3. Compare and contrast geom_jitter() with geom_count().

```{r}

?geom_jitter
?geom_count

```

```{r}

mpg %>% 
  ggplot(aes(cty, hwy)) +
  geom_count()

```

So geom_count() increases the area of points with more individual points there, unlike geom_jitter() which randomly scatters points slightly.

4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
```{r}

?geom_boxplot

```

position = "dodge2"

```{r}

mpg %>% 
  filter(manufacturer %in% c("dodge", "toyota", "volkswagen", "ford", "chevrolet")) %>%  
  ggplot(aes(manufacturer, displ)) +
  geom_boxplot() +
  coord_flip()

```

----------

# 3.9- Coordinate systems

coord_flip() is great when wanting to show data horizontally instead of vertically
```{r}

mpg %>% 
  ggplot(aes(class, hwy)) +
  geom_boxplot() +
  coord_flip()

```

coord_quickmap() sets the aspect ratio for maps
```{r}

nz <- map_data("nz")

nz %>% 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black")

```

```{r}

nz <- map_data("nz")

nz %>% 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  coord_quickmap()

```

coord_polar plots in polor coordinates

## Exercises

1. Turn a stacked bar chart into a pie chart using coord_polar().
```{r}

mpg %>% 
  ggplot(aes(cyl, fill = trans)) +
  geom_bar() +
  coord_polar()

```

2. What does labs() do? Read the documentation.
```{r}

mpg %>% 
  ggplot(aes(cyl, fill = trans)) +
  geom_bar() +
  coord_polar() +
  labs(title = "A stupid ass pie chart",
       subtitle = "So stupid",
       x = "Cylinders",
       y = "Count")

```

3. What’s the difference between coord_quickmap() and coord_map()?
```{r}

?coord_map

```

coord_quickmap() is like coord_map() but requires much less computation because it keeps all lines straight.

4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()

```

geom_abline() plots a custom line with a slope/intercept. On this plot, the points are above slope = 2 telling us that for an increase of 1 in cty, hwy is usually up 2+.

----------


